---
title: "FP 2"
author: ""
date: "April 24, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}

library(tidyverse)
library(fiftystater)
library(gridExtra)
library(sf)
library(ggmap)

```

##Names

Lynda Chao, Nick McMullen, Justin Reis-Henrie, Hannah Shumway

##Workflow

- Hannah and Lynda: working with shapefiles and visualizations in GIS
- Nick and Justin: working with visualizations and CSV data in RStudio
- All together: compiling work into shared Google Drive poster

- Meetings: Tuesday/Thursday at 11:30 AM

- The shared RMD will be compiled in a Github repository

##Topic and Research Questions

\textbf{Topic:} Gerrymandering Congressional Districts

\textbf{Research Questions:}
- Which states are, on average, most gerrymandered?
- Which states contain the top 5 most gerrymandered congressional districts?
- (Potential third research question) Which states contain the top 5 most gerrymandered state legislature districts and which party controls those state legislatures?


##Data

1. Average district gerrymandering (compactness and spread) metrics 

- Columns: State, Number of Districts, Polsby-Popper, Schwartzberg, Convex-Hull, Reock
- Limitations: data is on average by state and not on the Congressional District level
- Dimensions: 43 observations of 6 variables (did some data cleaning to add in states that only have one Congressional District)
- Most recent data for Congressional Districts

source: http://www.governing.com/gov-data/politics/gerrymandered-congressional-districts-compactness-by-state.html

```{r}

stateGerrymander <- read.csv("Data/StateGerrymandering.csv")

stateGerrymander <- stateGerrymander %>%
  mutate(Polsby.Popper = as.numeric(Polsby.Popper/100)) %>%
  mutate(Schwartzberg = as.numeric(Schwartzberg/100)) %>%
  mutate(Convex.Hull = as.numeric(Convex.Hull/100)) %>%
  mutate(Reock = as.numeric(Reock/100)) 


stateGerrymander$State <- as.character(stateGerrymander$State)

stateGerrymander <- stateGerrymander %>%
  rbind(c("Alaska", 1, 1, 1, 1, 1)) %>%
  rbind(c("Montana", 1, 1, 1, 1, 1)) %>%
  rbind(c("Delaware", 1, 1, 1, 1, 1)) %>%
  rbind(c("North Dakota", 1, 1, 1, 1, 1)) %>%
  rbind(c("South Dakota", 1, 1, 1, 1, 1)) %>%
  rbind(c("Vermont", 1, 1, 1, 1, 1)) %>%
  rbind(c("Wyoming", 1, 1, 1, 1, 1)) 

stateGerrymander$statelower <- tolower(stateGerrymander$State)

stateGerrymander$Polsby.Popper <- as.numeric(stateGerrymander$Polsby.Popper)
stateGerrymander$Schwartzberg <- as.numeric(stateGerrymander$Schwartzberg)
stateGerrymander$Convex.Hull <- as.numeric(stateGerrymander$Convex.Hull)
stateGerrymander$Reock <- as.numeric(stateGerrymander$Reock)

```


2. Congressional District Shapefile for 115th Congress 

- Not traditional data: we are working to attach gerrymander indices to the shapefile and visualize that way 
- Most recent data for Congressional Districts
- Source: US Census (census.gov)


3. Congressional District Compactness Data:

- PROBLEM: Doesn't have StateID or DistrictID to join with other data.

```{r}

cdCompact <- read.csv("https://raw.githubusercontent.com/jeffreyshen19/Is-My-District-Gerrymandered/master/data/master.csv")

```


4. Congressional District Raw Data - Area and Perimeter (but does not have comparable units with area and perimeter)

```{r}

congressDistrictRaw <- read.csv("Data/congressDistrictDataRaw.csv")

#Remove rows that are not congressional districts

congressDistrictRaw <- congressDistrictRaw %>%
  filter(!is.na(district_name))

#Calculate important metrics - unable to calculate Convex Hull and Reock w/o GIS - also, area and perimeter are in different units -- not sure which units -- this makes Polsby Popper calculations inaccurate, but Schwartzberg relies on a mere ratio, which we can use. 

congressDistrictRaw <- congressDistrictRaw %>%
  mutate(Schwartzberg = (2*pi*sqrt(area/pi)) / (perimeter))

```


5. State Legislative Districts 

- Columns: STATEFP, State, SLDLST, GEOID, NAMELSAD, Area, Perimeter, Polsby.Popper, Schwartzberg, Area.Convex.Hull, Reock
- Limitations: data is from 2014 not from most recent redraw of state legislature districts 
- Dimensions: 4825 observations 
- Source: https://raw.githubusercontent.com/cicero-data/compactness-stats/master/2014-us/state_lower.csv

```{r}

districtGerrymander <- read.csv("Data/districtGerrymander.csv")

```


##Initial Visualizations

Research Question 1: State level gerrymandering indices


```{r}

ggplot(stateGerrymander, aes(x=State, y=Polsby.Popper))+ 
  geom_bar(stat="identity") + 
  theme(axis.text.x = element_text(angle=90))


polsby_popper_map <- ggplot(stateGerrymander, aes(map_id = statelower)) + 
  geom_map(aes(fill=Polsby.Popper), map=fifty_states) + 
  expand_limits(x=fifty_states$long, y=fifty_states$lat) + 
  coord_map() + 
  scale_x_continuous(breaks=NULL) + 
  scale_y_continuous(breaks=NULL) + labs(x = "", y= "") + 
  theme(legend.position="bottom", panel.background = element_blank()) + 
  labs(title="Mean Polsby-Popper Ratio")


```


Schwartzberg

```{r}

schwartzberg_map <- ggplot(stateGerrymander, aes(map_id = statelower)) + 
  geom_map(aes(fill=Schwartzberg), map=fifty_states) + 
  expand_limits(x=fifty_states$long, y=fifty_states$lat) + 
  coord_map() + 
  scale_x_continuous(breaks=NULL) + 
  scale_y_continuous(breaks=NULL) + labs(x = "", y= "") + 
  theme(legend.position="bottom", panel.background = element_blank()) + 
  labs(title="Mean Schwartzberg Ratio")


```


Convex Hull Map

```{r}

convex_hull_map <- ggplot(stateGerrymander, aes(map_id = statelower)) + 
  geom_map(aes(fill=Convex.Hull), map=fifty_states) + 
  expand_limits(x=fifty_states$long, y=fifty_states$lat) + 
  coord_map() + 
  scale_x_continuous(breaks=NULL) + 
  scale_y_continuous(breaks=NULL) + labs(x = "", y= "") + 
  theme(legend.position="bottom", panel.background = element_blank()) + 
  labs(title="Mean Convex-Hull Ratio")


```


Reock Ratio Map

```{r}

reock_map <- ggplot(stateGerrymander, aes(map_id = statelower)) + 
  geom_map(aes(fill=Reock), map=fifty_states) + 
  expand_limits(x=fifty_states$long, y=fifty_states$lat) + 
  coord_map() + 
  scale_x_continuous(breaks=NULL) + 
  scale_y_continuous(breaks=NULL) + labs(x = "", y= "") + 
  theme(legend.position="bottom", panel.background = element_blank()) + 
  labs(title="Mean Reock Ratio")

```


All 4 plots together

```{r}

grid.arrange(polsby_popper_map, schwartzberg_map, convex_hull_map, reock_map, nrow=2, bottom="Figure 1")

```


Comparing 5 worst states for each metric:

```{r}

stateGerrymander %>%
  arrange(Polsby.Popper) %>%
  head(5)

stateGerrymander %>%
  arrange(Schwartzberg) %>%
  head(5)

stateGerrymander %>%
  arrange(Convex.Hull) %>%
  head(5)

stateGerrymander %>%
  arrange(Reock) %>%
  head(5)

```


We observe the 5 worst states in Polsby.Popper to be the same as the Schwartzberg ratio. Convex-Hull and Reock, both measures of spread, differ in the worst 5 significantly. 

Research Question 2: Top 5 Gerrymandered Congressional Districts


```{r}

Worst10 <- congressDistrictRaw %>%
  arrange(Schwartzberg) %>%
  head(10)

Worst10

```


States to investigate further and visualize:

- North Carolina
- Florida
- Maryland 
- Texas
- Pennsylvania
- Louisiana




