---
title: "Preliminary Data Analysis"
author: "Nick McMullen"
date: "April 11, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

library(tidyverse)
library(fiftystater)
library(gridExtra)
library(sf)
library(ggmap)

data("fifty_states")

stateGerrymander <- read.csv("Data/StateGerrymandering.csv")

```

Data from: http://www.governing.com/gov-data/politics/gerrymandered-congressional-districts-compactness-by-state.html

The stateGerrymander data imported above has information on the average compactness and spread metrics for each U.S. state that has more than one congressional district. 


Explore:

1. Compactness

- Polsby-Popper: this measure is a ratio of the area of the district to the area of the circle whose circumference is equal to the perimeter of the district. 


- Schwartzberg: this measure is a ratio of the perimeter of the district to the circumference of the circle whose area is equal to the area of the district. 

2. Spread

- Area/Convex Hull: this measure is a ratio of the area of the district to the area of the minimum convex polygod that can enclose the district.

- Reock: this measure is the ratio of the area of the district to the area of the minimum spanning circle that can enclose the district. 


##Analysis on the State Level 

NOTE: States with only one Congressional district are awarded ratios of 1 because the geometry of their districts are determined by state lines and not by politicians or independent commissions drawing boundaries. 

Clean the data:

```{r}

stateGerrymander <- stateGerrymander %>%
  mutate(Polsby.Popper = as.numeric(Polsby.Popper/100)) %>%
  mutate(Schwartzberg = as.numeric(Schwartzberg/100)) %>%
  mutate(Convex.Hull = as.numeric(Convex.Hull/100)) %>%
  mutate(Reock = as.numeric(Reock/100)) 


stateGerrymander$State <- as.character(stateGerrymander$State)

stateGerrymander <- stateGerrymander %>%
  rbind(c("Alaska", 1, 1, 1, 1, 1)) %>%
  rbind(c("Montana", 1, 1, 1, 1, 1)) %>%
  rbind(c("Delaware", 1, 1, 1, 1, 1)) %>%
  rbind(c("North Dakota", 1, 1, 1, 1, 1)) %>%
  rbind(c("South Dakota", 1, 1, 1, 1, 1)) %>%
  rbind(c("Vermont", 1, 1, 1, 1, 1)) %>%
  rbind(c("Wyoming", 1, 1, 1, 1, 1)) 

stateGerrymander$statelower <- tolower(stateGerrymander$State)

stateGerrymander$Polsby.Popper <- as.numeric(stateGerrymander$Polsby.Popper)
stateGerrymander$Schwartzberg <- as.numeric(stateGerrymander$Schwartzberg)
stateGerrymander$Convex.Hull <- as.numeric(stateGerrymander$Convex.Hull)
stateGerrymander$Reock <- as.numeric(stateGerrymander$Reock)



```


Exploratory plot:

```{r}

ggplot(stateGerrymander, aes(x=State, y=Polsby.Popper))+ 
  geom_bar(stat="identity") + 
  theme(axis.text.x = element_text(angle=90))


polsby_popper_map <- ggplot(stateGerrymander, aes(map_id = statelower)) + 
  geom_map(aes(fill=Polsby.Popper), map=fifty_states) + 
  expand_limits(x=fifty_states$long, y=fifty_states$lat) + 
  coord_map() + 
  scale_x_continuous(breaks=NULL) + 
  scale_y_continuous(breaks=NULL) + labs(x = "", y= "") + 
  theme(legend.position="bottom", panel.background = element_blank()) + 
  labs(title="Mean Polsby-Popper Ratio By State")

polsby_popper_map


```


Schwartzberg

```{r}

schwartzberg_map <- ggplot(stateGerrymander, aes(map_id = statelower)) + 
  geom_map(aes(fill=Schwartzberg), map=fifty_states) + 
  expand_limits(x=fifty_states$long, y=fifty_states$lat) + 
  coord_map() + 
  scale_x_continuous(breaks=NULL) + 
  scale_y_continuous(breaks=NULL) + labs(x = "", y= "") + 
  theme(legend.position="bottom", panel.background = element_blank()) + 
  labs(title="Mean Schwartzberg Ratio By State")

schwartzberg_map

```


Convex Hull Map

```{r}

convex_hull_map <- ggplot(stateGerrymander, aes(map_id = statelower)) + 
  geom_map(aes(fill=Convex.Hull), map=fifty_states) + 
  expand_limits(x=fifty_states$long, y=fifty_states$lat) + 
  coord_map() + 
  scale_x_continuous(breaks=NULL) + 
  scale_y_continuous(breaks=NULL) + labs(x = "", y= "") + 
  theme(legend.position="bottom", panel.background = element_blank()) + 
  labs(title="Mean Convex-Hull Ratio By State")

convex_hull_map

```


Reock Ratio Map

```{r}

reock_map <- ggplot(stateGerrymander, aes(map_id = statelower)) + 
  geom_map(aes(fill=Reock), map=fifty_states) + 
  expand_limits(x=fifty_states$long, y=fifty_states$lat) + 
  coord_map() + 
  scale_x_continuous(breaks=NULL) + 
  scale_y_continuous(breaks=NULL) + labs(x = "", y= "") + 
  theme(legend.position="bottom", panel.background = element_blank()) + 
  labs(title="Mean Reock Ratio By State")

reock_map

```


All 4 plots together

```{r}

grid.arrange(polsby_popper_map, schwartzberg_map, convex_hull_map, reock_map, nrow=2)

```



Notice that darker states have lower ratios, which indicates either a lack of compactness or a lack of even spread. Thus, darker states are states that are likely to have more gerrymandered districts. 

What are the 5 states with the worst average metrics?

Polsby-Popper

```{r}

stateGerrymander %>%
  arrange(Polsby.Popper) %>%
  select(State, Polsby.Popper) %>%
  head(5)

```

Schwartzberg

```{r}

stateGerrymander %>%
  arrange(Schwartzberg) %>%
  select(State, Schwartzberg) %>%
  head(5)

```

Convex-Hull

```{r}

stateGerrymander %>%
  arrange(Convex.Hull) %>%
  select(State, Convex.Hull) %>%
  head(5)

```

Reock

```{r}

stateGerrymander %>%
  arrange(Reock) %>%
  select(State, Reock) %>%
  head(5)

```


Notice that the first two metrics list Hawaii as the second worst gerrymandered state on average, even though it only has two districts. This is because the first two metrics don't do a good job accounting for oddly shaped state borders, including coastlines. Reock lists New Hampshire as the most gerrymandered state on average even though it only has 2 districts as well. 

Convex-Hull seems to do the best job, at least on the surface through this initial analysis. We should investigate these metrics further at a district level, with a special focus on the Convex-Hull Ratio. 


##Analysis on Congressional District Level

Let's look at the map on a district level:

Get the Congressional district maps

Map data from: http://cdmaps.polisci.ucla.edu/shp/districts%03i.zip

http://cdmaps.polisci.ucla.edu/tut/mapping_congress_in_R.html


```{r}

get_congress_map <- function(cong=113) {
  tmp_file <- tempfile()
  tmp_dir  <- tempdir()
  zp <- sprintf("http://cdmaps.polisci.ucla.edu/shp/districts%03i.zip",cong)
  download.file(zp, tmp_file)
  unzip(zipfile = tmp_file, exdir = tmp_dir)
  fpath <- paste(tmp_dir, sprintf("districtShapes/districts%03i.shp",cong), sep = "/")
  st_read(fpath)
}

```

Then we grab data from the 114th Congress (2015-2017):

```{r}

cd114 <- get_congress_map(114)

```


```{r}

congressDistrictGerrymander <- read.csv("Data/congressDistrictGerrymander.csv")

congressDistrictGerrymander <- congressDistrictGerrymander %>%
  mutate(Polsby.Popper = compactness) %>%
  select(district, Polsby.Popper) 

congressDistrictGerrymander$Polsby.Popper[1] = 1
congressDistrictGerrymander$Polsby.Popper[87] = 1
congressDistrictGerrymander$Polsby.Popper[233] = 1
congressDistrictGerrymander$Polsby.Popper[247] = 1
congressDistrictGerrymander$Polsby.Popper[352] = 1
congressDistrictGerrymander$Polsby.Popper[413] = 1
congressDistrictGerrymander$Polsby.Popper[435] = 1


```

Source: https://raw.githubusercontent.com/jeffreyshen19/Is-My-District-Gerrymandered/master/data/compactness.csv


Look at the 5 worst districts for Polsby-Popper:

```{r}

congressDistrictGerrymander %>%
  arrange(Polsby.Popper) %>%
  head(10)

```



##Analysis on State House District Level


Let's import some data that contains information on the State House district level:

```{r}

districtGerrymander <- read.csv("Data/districtGerrymander.csv")

```

Data from: https://raw.githubusercontent.com/cicero-data/compactness-stats/master/2014-us/state_lower.csv

This data has all of the information and calculations for all four metrics (averages) we described above for each state. 













